---
title: "Comp Stats | Give Me Some Credit"
author: "Lathan Liou, Shivang Mehta, Isaac Cui, Abby Lewis"
date: "11/30/2018"
output: pdf_document
---

```{r setup, include=FALSE message = FALSE}
## Section 0: Environment Setup

knitr::opts_chunk$set(echo = TRUE)
#### Libraries ####
library(readxl)
library(tidyr)
library(VIM)
library(readr)
library(dplyr)
library(glmnet)
library(ggplot2)
library(mice)
library(scales)
library(DescTools)
library(randomForest)
library(pROC)
library(data.table)
library(xgboost)


cs_training <- read_csv("C:/Users/Shivang Mehta/Desktop/Work/Fall 18/Computational Statistics/Project/Data/cs-training.xlsx")

#remove first column
cs_training <- cs_training[,-1]

#making all columns numeric
cs_training <- as.data.frame(cs_training)
```
\par 
Banks play a crucial role in market economies. They decide who can get finance and on what terms and can make or break investment decisions. For markets and society to function, individuals and companies need access to credit. 
\par 
Credit scoring algorithms, which make a guess at the probability of default, are the method banks use to determine whether or not a loan should be granted. Through this project, we attempt to improve on the state of the art in credit scoring, by predicting the probability that somebody will experience financial distress in the next two years.
\par 
We begin by exploring the data provided to us:
\par 
```{r}
## Section 1: Data Exploration

# Dependent variable 
ggplot(cs_training, aes(SeriousDlqin2yrs)) + geom_bar(aes(y= (..count..)/sum(..count..)), fill = "steelblue", color ="steelblue") + ylab("Percent")

# Independent Variables

a1 <- ggplot(cs_training, aes(RevolvingUtilizationOfUnsecuredLines)) + geom_bar(aes(y= (..count..)/sum(..count..)), fill = "steelblue", color ="steelblue") + ylab("Percent")
a2 <- ggplot(cs_training, aes(age)) + geom_bar(fill = "steelblue", color ="steelblue")
a3 <- ggplot(cs_training, aes(`NumberOfTime30-59DaysPastDueNotWorse`)) + geom_bar(aes(y= (..count..)/sum(..count..)), fill = "steelblue", color ="steelblue") + ylab("Percent")
a4 <- ggplot(cs_training, aes(DebtRatio)) + geom_bar(fill = "steelblue", color ="steelblue") + ylab("Percent")

a5 <- ggplot(cs_training, aes(MonthlyIncome)) + geom_bar(fill = "steelblue", color ="steelblue") + ylab("Percent")
a6 <- ggplot(cs_training, aes(NumberOfOpenCreditLinesAndLoans)) + geom_bar(fill = "steelblue", color ="steelblue") + ylab("Percent")
a7 <- ggplot(cs_training, aes(`NumberOfTimes90DaysLate`)) + geom_bar(aes(y= (..count..)/sum(..count..)), fill = "steelblue", color ="steelblue") + ylab("Percent")
a8 <- ggplot(cs_training, aes(NumberRealEstateLoansOrLines)) + geom_bar(aes(y= (..count..)/sum(..count..)), fill = "steelblue", color ="steelblue") + ylab("Percent")

a9 <- ggplot(cs_training, aes(NumberOfDependents)) + geom_bar(aes(y= (..count..)/sum(..count..)), fill = "steelblue", color ="steelblue") + ylab("Percent")
a10 <- ggplot(cs_training, aes(`NumberOfTime60-89DaysPastDueNotWorse`)) + geom_bar(aes(y= (..count..)/sum(..count..)), fill = "steelblue", color ="steelblue") + ylab("Percent")
```




```{r}
## Section 2: Data Imputation

#EDA
md.pattern(cs_training)
aggr_plot <- aggr(cs_training, col=c('dodgerblue','black'), numbers=TRUE, sortVars=TRUE, labels=names(cs_training), cex.axis=.4, gap=0.5, ylab=c("Histogram of missing data","Pattern"))
```

NOTE ON USE OF MEDIAN FOR IMPUTATION AND REGRESSION.

# Imputing NumberOfDependents as median

```{r}
median.numdep <- median(cs_training$NumberOfDependents, na.rm = TRUE)
for(i in 1:length(cs_training$NumberOfDependents)){
  if(is.na(cs_training$NumberOfDependents[i])){
    cs_training$NumberOfDependents[i] = median.numdep
  }
}
```

# Imputing MonthlyIncome

```{r}
train1 <- cs_training
train1 <- train1 %>% drop_na()
train.y <- train1$MonthlyIncome
train1 <- subset(train1, select = -c(MonthlyIncome, SeriousDlqin2yrs))
```

# Regression

```{r}
fit <- lm(train.y ~., data=train1)
pred.miss <- fit$coefficients[1] + fit$coefficients[2]*cs_training$RevolvingUtilizationOfUnsecuredLines + fit$coefficients[3]*cs_training$age +
              fit$coefficients[4]*cs_training$`NumberOfTime30-59DaysPastDueNotWorse` + fit$coefficients[5]*cs_training$DebtRatio +
              fit$coefficients[6]*cs_training$NumberOfOpenCreditLinesAndLoans + fit$coefficients[7]*cs_training$NumberOfTimes90DaysLate +
              fit$coefficients[8]*cs_training$NumberOfOpenCreditLinesAndLoans + fit$coefficients[9]*cs_training$`NumberOfTime60-89DaysPastDueNotWorse` +
              fit$coefficients[10]*cs_training$NumberOfDependents
for(i in 1:length(cs_training$MonthlyIncome)){
  if(is.na(cs_training$MonthlyIncome[i])){
    cs_training$MonthlyIncome[i] = pred.miss[i]
  }
}

```

## Section 3: Outlier Treatment

SHORT NOTE ON WINSORIZATION

```{r}
## Winsorize Data 
cs_training_winsorized <- sapply(cs_training, Winsorize, minval = NULL)
cs_training_winsorized <- as.data.frame(cs_training_winsorized)
```
Maybe write a brief description of what the code did.

## Section 4: Feature Engineering

In class, we learned how adding basis functions could improve the way we understand our data: namely finding our decision rules in higher dimensions. Thus, feature engineering is a crucial part of the machine learning process. By transforming our data into features that we think better represent the underlying problem for our machine learning models, we hope to improve predictive performance. We tried to use our domain knowledge about credit scores and our creativity to come up with new features. We ultimately came up with 14 new features in addition to the existing 10 explanatory ones in the data set. 

```{r}
#add features
cs_training_winsorized <- cs_training_winsorized %>%
  mutate(Distributedincome = MonthlyIncome/(1+NumberOfDependents),
         Totalcost = DebtRatio * MonthlyIncome,
         Retired = ifelse(age > 65, 1, 0),
         Lateindex = `NumberOfTime30-59DaysPastDueNotWorse` + 2*`NumberOfTime60-89DaysPastDueNotWorse` + 3*NumberOfTimes90DaysLate,
         Numedepencubed = NumberOfDependents^3,
         LogRUUL = log(RevolvingUtilizationOfUnsecuredLines+1),
         LogMI = log(MonthlyIncome+1),
         LoansAge = (NumberOfOpenCreditLinesAndLoans + NumberRealEstateLoansOrLines)/age,
         DepAge = NumberOfDependents/age,
         LoansIncome = NumberOfOpenCreditLinesAndLoans + NumberRealEstateLoansOrLines,
         RUULIncome = log((RevolvingUtilizationOfUnsecuredLines+1)/(MonthlyIncome+1)), #Housing Insolvency paper
         HousingExpenses = log((1+NumberRealEstateLoansOrLines)/(MonthlyIncome+1)),
         AgeCubed = age^3,
         LogAge = log(age)
)

write_csv(cs_training_winsorized, "cs_training_clean.csv")
cs_training_clean <- read_csv("cs_training_clean.csv")


```

```{r}
ggplot(cs_training_clean, aes(LogAge, fill = as.factor(SeriousDlqin2yrs))) + geom_density(alpha = 0.2)
```

## Section 5: Feature Selection

```{r}
#xmat <- as.matrix(cs_training_clean[,-1])
#lassovar <- cv.glmnet(xmat, y=as.factor(cs_training_clean$SeriousDlqin2yrs), alpha = 0.5, lambda = seq(0.00051,0.00058,length.out=7),family="binomial")

```

## Section 6: Loading Test Data 

```{r}

cs_test <- cs_test[,-1]
cs_test <- cs_test[,-1]

# Imputation
median.numdep <- median(cs_test$NumberOfDependents, na.rm = TRUE)
for(i in 1:length(cs_test$NumberOfDependents)){
  if(is.na(cs_test$NumberOfDependents[i])){
    cs_test$NumberOfDependents[i] = median.numdep
  }
}

train1 <- cs_test
train1 <- train1 %>% drop_na()
train.y <- train1$MonthlyIncome
train1 <- subset(train1, select = -c(MonthlyIncome))


fit <- lm(train.y ~., data=train1)
pred.miss <- fit$coefficients[1] + fit$coefficients[2]*cs_test$RevolvingUtilizationOfUnsecuredLines + fit$coefficients[3]*cs_test$age +
              fit$coefficients[4]*cs_test$`NumberOfTime30-59DaysPastDueNotWorse` + fit$coefficients[5]*cs_test$DebtRatio +
              fit$coefficients[6]*cs_test$NumberOfOpenCreditLinesAndLoans + fit$coefficients[7]*cs_test$NumberOfTimes90DaysLate +
              fit$coefficients[8]*cs_test$NumberOfOpenCreditLinesAndLoans + fit$coefficients[9]*cs_test$`NumberOfTime60-89DaysPastDueNotWorse` +
              fit$coefficients[10]*cs_test$NumberOfDependents
for(i in 1:length(cs_test$MonthlyIncome)){
  if(is.na(cs_test$MonthlyIncome[i])){
    cs_test$MonthlyIncome[i] = pred.miss[i]
  }
}

# Winsorization
cs_test_winsorized <- sapply(cs_test, Winsorize, minval = NULL)
cs_test_winsorized <- as.data.frame(cs_test_winsorized)

# Feature Engineering 
cs_test_winsorized <- cs_test_winsorized %>%
  mutate(Distributedincome = MonthlyIncome/(1+NumberOfDependents),
         Totalcost = DebtRatio * MonthlyIncome,
         Retired = ifelse(age > 65, 1, 0),
         Lateindex = `NumberOfTime30-59DaysPastDueNotWorse` + 2*`NumberOfTime60-89DaysPastDueNotWorse` + 3*NumberOfTimes90DaysLate,
         Numedepencubed = NumberOfDependents^3,
         LogRUUL = log(RevolvingUtilizationOfUnsecuredLines+1),
         LogMI = log(MonthlyIncome+1),
         LoansAge = (NumberOfOpenCreditLinesAndLoans + NumberRealEstateLoansOrLines)/age,
         DepAge = NumberOfDependents/age,
         LoansIncome = NumberOfOpenCreditLinesAndLoans + NumberRealEstateLoansOrLines,
         RUULIncome = log((RevolvingUtilizationOfUnsecuredLines+1)/(MonthlyIncome+1)), #Housing Insolvency paper
         HousingExpenses = log((1+NumberRealEstateLoansOrLines)/(MonthlyIncome+1)),
         AgeCubed = age^3,
         LogAge = log(age)
)

write_csv(cs_test_winsorized, "cs_test_clean.csv")

```

## Section 7: Predictive Algorithm

```{r}
cs_test_clean <- read_csv("cs_test_clean.csv")
cs_training_clean <- read_csv("cs_training_clean.csv")

## Random Forest

output.forest<- randomForest(as.factor(cs_training_clean$SeriousDlqin2yrs) ~ cs_training_clean$RevolvingUtilizationOfUnsecuredLines+ cs_training_clean$age+ cs_training_clean$`NumberOfTime30-59DaysPastDueNotWorse`+ cs_training_clean$DebtRatio + cs_training_clean$MonthlyIncome + cs_training_clean$NumberOfOpenCreditLinesAndLoans + cs_training_clean$NumberOfTimes90DaysLate + cs_training_clean$NumberRealEstateLoansOrLines + cs_training_clean$`NumberOfTime60-89DaysPastDueNotWorse` + cs_training_clean$NumberOfDependents + cs_training_clean$Distributedincome + cs_training_clean$Totalcost + cs_training_clean$Retired + cs_training_clean$Lateindex + cs_training_clean$Numedepencubed + cs_training_clean$LogRUUL + cs_training_clean$LogMI + cs_training_clean$LoansAge +cs_training_clean$DepAge + cs_training_clean$LoansIncome + cs_training_clean$RUULIncome + cs_training_clean$HousingExpenses + cs_training_clean$AgeCubed + cs_training_clean$LogAge , 
                             data=cs_training_clean, ntree=100, importance=T, replace=T, xtest = cs_test_winsorized)

y.test <- output.forest$test
```

```{r}
#### xgboost ####

#convert data frame to matrix
cs_xgboost <- cs_training_clean[,-1]
cs_xgboost <- as.matrix(cs_xgboost)

cs_test_clean <- as.matrix(cs_test_clean)

creditboost <- xgboost(data = cs_xgboost, max.depth = 5, label = as.matrix(cs_training_clean[,1]),
               eta = 0.1, nthread = 2, nround = 2, objective = "binary:logistic")

pred <- predict(creditboost, cs_test_clean)

submission.xgboost <- pred
submission.xgboost <- as.data.frame(submission.xgboost)
id <- seq(1,nrow(submission.xgboost), by = 1)
submission.xgboost <- cbind(as.integer(id), submission.xgboost)
colnames(submission.xgboost) <- c("id", "Probability")
write.csv(submission.xgboost, "submission.xgboost.csv")

```

## Section 8: Output file
```{r}
submissions <- output.forest$test$votes[,2]
submissions <- as.data.frame(submission)
id <- seq(1,nrow(submission), by = 1)
submissions <- cbind(id, submissions)
colnames(submissions) <- c("id", "Probability")
write.csv(submissions, "submissions.csv")
```

## Section 9: Trying xgboost's imputation method
```{r}
#prepare training data
cs_training_a <- sapply(cs_training, Winsorize, minval = NULL, na.rm = TRUE)
cs_training_a <- as.data.frame(cs_training_a)

# Feature Engineering 
cs_training_a <- cs_training_a %>%
  mutate(Distributedincome = MonthlyIncome/(1+NumberOfDependents),
         Totalcost = DebtRatio * MonthlyIncome,
         Retired = ifelse(age > 65, 1, 0),
         Lateindex = `NumberOfTime30-59DaysPastDueNotWorse` + 2*`NumberOfTime60-89DaysPastDueNotWorse` + 3*NumberOfTimes90DaysLate,
         Numedepencubed = NumberOfDependents^3,
         LogRUUL = log(RevolvingUtilizationOfUnsecuredLines+1),
         LogMI = log(MonthlyIncome+1),
         LoansAge = (NumberOfOpenCreditLinesAndLoans + NumberRealEstateLoansOrLines)/age,
         DepAge = NumberOfDependents/age,
         LoansIncome = NumberOfOpenCreditLinesAndLoans + NumberRealEstateLoansOrLines,
         RUULIncome = log((RevolvingUtilizationOfUnsecuredLines+1)/(MonthlyIncome+1)), #Housing Insolvency paper
         HousingExpenses = log((1+NumberRealEstateLoansOrLines)/(MonthlyIncome+1)),
         AgeCubed = age^3,
         LogAge = log(age)
)

cs_xgboost_a <- cs_training_a[,-c(1:2)]
cs_xgboost_a <- as.matrix(cs_xgboost_a)

creditboost <- xgboost(data = cs_xgboost_a, max.depth = 5, label = cs_xgboost_a[,1],
               eta = 1, nthread = 2, nround = 2, objective = "binary:logistic", missing = NA)

#prepare test data
cs_test_a <- sapply(cs_test, Winsorize, minval = NULL, na.rm = TRUE)
cs_test_a <- as.data.frame(cs_test_a)

# Feature Engineering 
cs_test_a <- cs_test_a %>%
  mutate(Distributedincome = MonthlyIncome/(1+NumberOfDependents),
         Totalcost = DebtRatio * MonthlyIncome,
         Retired = ifelse(age > 65, 1, 0),
         Lateindex = `NumberOfTime30-59DaysPastDueNotWorse` + 2*`NumberOfTime60-89DaysPastDueNotWorse` + 3*NumberOfTimes90DaysLate,
         Numedepencubed = NumberOfDependents^3,
         LogRUUL = log(RevolvingUtilizationOfUnsecuredLines+1),
         LogMI = log(MonthlyIncome+1),
         LoansAge = (NumberOfOpenCreditLinesAndLoans + NumberRealEstateLoansOrLines)/age,
         DepAge = NumberOfDependents/age,
         LoansIncome = NumberOfOpenCreditLinesAndLoans + NumberRealEstateLoansOrLines,
         RUULIncome = log((RevolvingUtilizationOfUnsecuredLines+1)/(MonthlyIncome+1)), #Housing Insolvency paper
         HousingExpenses = log((1+NumberRealEstateLoansOrLines)/(MonthlyIncome+1)),
         AgeCubed = age^3,
         LogAge = log(age)
)

cs_test_a <- cs_test_a[,-c(1:2)]
cs_test_a <- as.matrix(cs_test_a)

pred <- predict(creditboost, cs_test_a)

submission.xgboost <- pred
submission.xgboost <- as.data.frame(submission.xgboost)
id <- seq(1,nrow(submission.xgboost), by = 1)
submission.xgboost <- cbind(as.integer(id), submission.xgboost)
colnames(submission.xgboost) <- c("id", "Probability")
write.csv(submission.xgboost, "submission.xgboost2.csv")
```
